---
title: "Meta-analyses of item content similarity in clinical domains"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# TODO

other domains:

- AUD to be added
- schizophrenia/psychosis: not enough self report symptom scales when Tim looked into it.
- BPD?

```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(dplyr)
library(readr)
library(tibble)
library(forcats)
library(proxy)
library(janitor)
library(metafor)
library(ggplot2)
library(ggstance)
library(scales)
library(knitr)
library(kableExtra)

```

# Data

```{r}

data_depression <- read_csv("../data/depression - Fried 2017/MatrixB.csv")     #Load dat for estimating Jaccard index (no difference between specific and compound symptoms)

data_anxiety <- read_csv("../data/anxiety - Wall & Lee 2022/anxiety_wall_lee_jaccard.csv")

data_psychosis <- read_csv("../data/psychosis - Bernardin et al. 2023/psychosis - Bernardin et al. 2023.csv") |> 
  janitor::clean_names() %>%
  mutate(across(everything(), ~ ifelse(. == 2, 1, .)))

data_hypomania <- read_csv("../data/hypomania - Chrobak et al. 2018/hypomania.csv") 

data_eatingdisorders <- read_csv("../data/eating disorders - Christensen Pacella et al. 2024/MatrixB_EDMeasures_5-26-2023.csv")

data_youth_ocd <- read_csv("../data/ocd - Visiontay et al. 2019/youth_ocd.csv") |> 
  janitor::clean_names() %>%
  mutate(across(everything(), ~ ifelse(. == 2, 1, .)))

# youth depression

# youth anxiety

```

# Jaccard index interpretation

Real, R., & Vargas, J. M. (1996). The probabilistic basis of Jaccard's index of similarity. Systematic Biology, 45(3), 380-385. DOI:10.1093/sysbio/45.3.380

- Jaccard < 0.25: Typically considered low similarity, meaning that the two sets have few elements in common.
- 0.25 ≤ Jaccard < 0.5: Moderate similarity, which indicates that the sets share some commonalities but still have many differences.
- 0.5 ≤ Jaccard < 0.75: High similarity, meaning there is substantial overlap between the sets.
- Jaccard ≥ 0.75: Very high similarity, with most elements being shared between the sets.

# Mean Jaccard index by domain

```{r}

cor_estimates <- function(.data){
  data_mat_t <- .data |>
    as.matrix() |> 
    t()
  
  jaccard_diff <- proxy::dist(data_mat_t, method = "Jaccard") |> as.matrix()
  jaccard_sim <- 1 - jaccard_diff
  
  jaccard_sim[!lower.tri(jaccard_sim)] <- NA
  
  mean_jaccard <- mean(jaccard_sim, na.rm = TRUE)
  se_jaccard <- plotrix::std.error(as.vector(jaccard_sim), na.rm = TRUE)
  
  return(data.frame(J_mean = mean_jaccard,
                    J_se = se_jaccard,
                    J_ci_lower = mean_jaccard - se_jaccard*1.96,
                    J_ci_upper = mean_jaccard + se_jaccard*1.96))
}

dat <- bind_rows(
  data_depression |>
    cor_estimates() |>
    mutate(domain = "Depression"),
  data_anxiety |>
    cor_estimates() |>
    mutate(domain = "Anxiety"),
  data_psychosis |>
    cor_estimates() |>
    mutate(domain = "Psychosis"),
  data_hypomania |>
    cor_estimates() |>
    mutate(domain = "(Hypo)mania"),
  data_youth_ocd |>
    cor_estimates() |>
    mutate(domain = "(Youth) OCD"),
  data_eatingdisorders |>
    cor_estimates() |>
    mutate(domain = "Eating Disorders")
) |>
  mutate(logit_J_mean = qlogis(J_mean),
         # Calculate SEs on the logit scale using the delta method
         # SE_logit_p = SE_p / [p * (1 - p)]
         logit_J_se = J_se / (J_mean * (1 - J_mean))) |>
  relocate(domain, .before = J_mean)

dat |>
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## Meta across domains

```{r}

# fit meta
fit <- rma(yi = logit_J_mean, 
           vi = logit_J_se^2,
           data = dat,
           method = "REML")

# create a forest plot
plot_forest <- 
  metafor::forest(fit, 
                  transf = plogis,
                  slab = domain,
                  xlim = c(-0.8, 1.7),
                  at = c(0, .2, .4, .6, .8, 1),
                  efac = 1,
                  addpred = TRUE,
                  xlab = "Jaccard Similarity",
                  header = c("Domain", "J [95% CI]"),
                  refline = NA)

# table of results
results_predictions <- 
  bind_cols(
    fit |>
      predict(transf = plogis, level = 50) |>
      as_tibble() |>
      dplyr::select(meta_J = pred, ci_lower = ci.lb, ci_upper = ci.ub, pi_lower_50 = pi.lb, pi_upper_50 = pi.ub),
    fit |>
      predict(transf = plogis, level = 95) |>
      as_tibble() |>
      dplyr::select(pi_lower_95 = pi.lb, pi_upper_95 = pi.ub)
  ) |>
  mutate_all(janitor::round_half_up, digits = 2)

results_predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

# 3-level meta-analyses by domain

```{r}

escalc_jaccard <- function(.data, domain = NA) {
  data_mat_t <- .data |> 
    as.matrix() |> 
    t()

  jaccard_difference <- proxy::dist(data_mat_t, method = "Jaccard") |> as.matrix()
  jaccard_similarity <- 1 - jaccard_difference

  # Calculate SEs for each Jaccard similarity
  var_names <- rownames(data_mat_t)
  n_vars <- length(var_names)

  # Initialize the SE matrix
  se_matrix <- matrix(NA, nrow = n_vars, ncol = n_vars)
  rownames(se_matrix) <- var_names
  colnames(se_matrix) <- var_names

  # Compute SEs
  for (i in 1:n_vars) {
    for (j in 1:n_vars) {
      x <- data_mat_t[i, ]
      y <- data_mat_t[j, ]
      a <- sum(x == 1 & y == 1)
      b <- sum(x == 1 & y == 0)
      c <- sum(x == 0 & y == 1)
      n <- a + b + c
      J <- a / n
      SE <- sqrt(J * (1 - J) / n)
      se_matrix[i, j] <- SE
    }
  }

  # Dynamic epsilon for zero-and-one adjustment
  #epsilon <- 1 / n_vars  # Proportional to the number of variables
  epsilon <- 0.5 # common default, greater pull towards center and less influence of 0,1 values
  
  # Apply zero-and-one adjustment
  adjusted_jaccard_similarity <- (jaccard_similarity * (n_vars - 1) + epsilon) / n_vars

  # Logit transform
  logit_jaccard_similarity <- qlogis(adjusted_jaccard_similarity)

  # Calculate SEs on the logit scale using the delta method
  # SE_logit_p = SE_p / [p * (1 - p)]
  se_logit_jaccard_similarity <- se_matrix / (adjusted_jaccard_similarity * (1 - adjusted_jaccard_similarity))

  # Replace zero SEs with a principled lower bound
  min_se <- sqrt(epsilon * (1 - epsilon) / n_vars)
  se_logit_jaccard_similarity[se_logit_jaccard_similarity == 0] <- min_se

  # Replace infinite values resulting from division by zero
  se_logit_jaccard_similarity[!is.finite(se_logit_jaccard_similarity)] <- NA

  # Create variable pair names
  m <- adjusted_jaccard_similarity
  ## Get indices and values
  indices <- which(lower.tri(m), arr.ind = TRUE)
  vec <- m[lower.tri(m)]
  ## Create variable pair names
  variable_pairs <- paste(rownames(m)[indices[, 1]], colnames(m)[indices[, 2]], sep = " - ")

  # Create a data frame for effect sizes
  dat_es <- tibble(
    domain = domain,
    J = jaccard_similarity[lower.tri(jaccard_similarity)],
    yi = logit_jaccard_similarity[lower.tri(logit_jaccard_similarity)],
    sei = se_logit_jaccard_similarity[lower.tri(se_logit_jaccard_similarity)],
    scale1 = rownames(m)[indices[, 1]],
    scale2 = colnames(m)[indices[, 2]],
    pair = variable_pairs
  )

  return(dat_es)
}


rma_jaccard_three_level <- function(effect_sizes, title = "Scales"){
  
  # fit a three-level meta-analysis model accounting for shared scales
  fit <- rma.mv(yi = yi, 
                V = sei^2,
                random = list(~ 1 | scale1, ~ 1 | scale2),
                data = effect_sizes,
                method = "REML")
  
  # create a forest plot
  plot_forest <- 
    metafor::forest(fit, 
                    transf = plogis,
                    slab = effect_sizes$pair,
                    xlim = c(-0.8, 1.7),
                    at = c(0, .2, .4, .6, .8, 1),
                    efac = 0.2,
                    addpred = TRUE,
                    xlab = "Jaccard Similarity",
                    header = c(title, "J [95% CI]"),
                    refline = NA)
  
  # table of results
  results_predictions <- 
    bind_cols(
      fit |>
        predict(transf = plogis, level = 50) |>
        as_tibble() |>
        dplyr::select(J = pred, ci_lower = ci.lb, ci_upper = ci.ub, pi_lower_50 = pi.lb, pi_upper_50 = pi.ub),
      fit |>
        predict(transf = plogis, level = 95) |>
        as_tibble() |>
        dplyr::select(pi_lower_95 = pi.lb, pi_upper_95 = pi.ub)
    ) |>
    mutate(domain = title) |>
    relocate(domain, .before = J)
  
  return(list(fit = fit,
              predictions = results_predictions,
              forest = plot_forest))
}

```

## Depression

```{r fig.height=6, fig.width=6}

es_depression <- data_depression |>
  escalc_jaccard(domain = "Depression") 

res_depression <- es_depression |>
  rma_jaccard_three_level(title = "Depression scales")

res_depression$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## Anxiety

```{r fig.height=16, fig.width=6}

es_anxiety <- data_anxiety |>
  escalc_jaccard(domain = "Anxiety") 

res_anxiety <- es_anxiety |>
  rma_jaccard_three_level(title = "Anxiety scales")

res_anxiety$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## Psychosis

```{r fig.height=16, fig.width=6}

es_psychosis <- data_psychosis |>
  escalc_jaccard(domain = "Psychosis")

res_psychosis <- es_psychosis |>
  rma_jaccard_three_level(title = "Psychosis scales")

res_psychosis$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## Hypomania

```{r fig.height=6, fig.width=6}

es_hypomania <- data_hypomania |>
  escalc_jaccard(domain = "(Hypo)mania")

res_hypomania <- es_hypomania |>
  rma_jaccard_three_level(title = "(Hypo)mania scales")

res_hypomania$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## OCD (youth)

```{r fig.height=8, fig.width=6}

es_youth_ocd <- data_youth_ocd |>
  escalc_jaccard(domain = "Youth OCD")

res_youth_ocd <- es_youth_ocd |>
  rma_jaccard_three_level(title = "Youth OCD scales")

res_youth_ocd$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## Eating disorders

```{r fig.height=5, fig.width=6}

es_eatingdisorders <- data_eatingdisorders |>
  escalc_jaccard(domain = "OCD")

res_eatingdisorders <- es_eatingdisorders |>
  rma_jaccard_three_level(title = "Eating disorders scales")

res_eatingdisorders$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

# 4-level meta-analyses across domains

```{r fig.height=33, fig.width=6}

es_combined <- bind_rows(
  es_depression,
  es_anxiety,
  es_psychosis,
  es_hypomania,
  es_eatingdisorders,
  es_youth_ocd
)

rma_jaccard_four_level <- function(effect_sizes){
  
  # Fit a four-level meta-analysis model accounting for shared scales
  fit <- rma.mv(yi = yi,
                V = sei^2,
                random = list(~ 1 | domain/scale1, ~ 1 | domain/scale2),
                data = effect_sizes,
                method = "REML")
  
  # Create a forest plot
  plot_forest <- 
    meta::forest(fit, 
           transf = plogis,
           slab = effect_sizes$pair,
           xlim = c(-0.8, 1.7),
           at = c(0, .2, .4, .6, .8, 1),
           #cex = 0.4,
           efac = 0.2,
           addpred = TRUE,
           xlab = "Jaccard Similarity",
           header = c("Scales", "J [95% CI]"),
           refline = NA)
  
  results_predictions <- 
    bind_cols(
      fit |>
        predict(transf = plogis, level = 50) |>
        as_tibble() |>
        dplyr::select(J = pred, ci_lower = ci.lb, ci_upper = ci.ub, pi_lower_50 = pi.lb, pi_upper_50 = pi.ub),
      fit |>
        predict(transf = plogis, level = 95) |>
        as_tibble() |>
        dplyr::select(pi_lower_95 = pi.lb, pi_upper_95 = pi.ub)
    ) |>
    mutate(domain = "4-level RE") |>
    relocate(domain, .before = J)
  
  return(list(fit = fit,
              predictions = results_predictions,
              forest = plot_forest))
}

res_combined <- rma_jaccard_four_level(es_combined)

res_combined$predictions |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

#res_combined$fit
 
# Combine estimates and labels into a data frame for easy viewing
tibble(label = c("domain", "domain/scale1", "domain again", "domain/scale2")) |>
  mutate(logit_sigma = res_combined$fit$sigma2,
         sigma = round_half_up(plogis(logit_sigma), 2)) |>
  filter(label != "domain again") |>
  select(-logit_sigma) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

```

## Comparions between domains

3- and 4-level meta

```{r fig.height=3, fig.width=6}

predictions_combined_by_domain <- bind_rows(
  res_depression$predictions,
  res_anxiety$predictions,
  res_psychosis$predictions,
  res_hypomania$predictions,
  res_youth_ocd$predictions,
  res_eatingdisorders$predictions,
  res_combined$predictions
) |>
  mutate(domain = fct_rev(fct_relevel(domain,
                                      "Depression scales",
                                      "Anxiety scales",
                                      "Psychosis scales",
                                      "(Hypo)mania scales",
                                      "Youth OCD scales",
                                      "Eating disorders scales",
                                      "4-level RE")))

predictions_combined_by_domain |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable(align = "r") |>
  kable_classic(full_width = FALSE)

ggplot(predictions_combined_by_domain, aes(J, domain)) +
  # geom_errorbarh(aes(xmin = pi_lower, xmax = pi_upper), linetype = "dotted", height = 0.26) +
  # geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), linetype = "solid", height = 0.26, size = 0.85) +
  geom_linerange(aes(xmin = pi_lower_95, xmax = pi_upper_95), linetype = "solid", linewidth = 0.3) +
  geom_linerange(aes(xmin = pi_lower_50, xmax = pi_upper_50), linetype = "solid", linewidth = 0.6, color = "red") +
  geom_linerange(aes(xmin = ci_lower, xmax = ci_upper), linetype = "solid", linewidth = 1.2) +
  geom_point(size = 2.4, color = "red") +
  theme_linedraw() +
  scale_x_continuous(limits = c(0,1), breaks = scales::breaks_pretty(n = 10), name = "Jaccard Similarity") +
  ylab("")

dir.create("plots")

ggsave("plots/plot.pdf",
       width = 6,
       height = 3)

```

# Session info

```{r}

sessionInfo()

```


