---
title: "Fried (2017) analyses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(tibble)
library(proxy)
library(vegan)
library(janitor)

library(qgraph)
library(ggplot2)
library(data.table)
library(reshape2)
library(psych)
library(ade4)
library(viridis)

```

# Data

```{r}

dat <- read_csv("MatrixB.csv")     #Load dat for estimating Jaccard index (no difference between specific and compound symptoms)

# dat <- tibble(
#   IDS   = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), 
#   QIDS  = c(1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), 
#   BDI   = c(1, 1, 1, 1, 1, 0, 0, 1, 0, 0, 0, 1, 1, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), 
#   CESD  = c(1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0), 
#   SDS   = c(1, 1, 1, 0, 1, 0, 0, 1, 0, 1, 0, 1, 1, 1, 0, 1, 1, 0, 0, 1, 1, 0, 1, 1, 0, 1, 0, 1, 0, 1, 1, 0, 0, 0, 0, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0), 
#   MADRS = c(1, 1, 1, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 1, 0, 0, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0), 
#   HRSD  = c(1, 1, 1, 0, 1, 1, 1, 1, 0, 0, 0, 1, 0, 1, 0, 0, 0, 1, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1)
# )

```

# Jaccard index

## Interpretation

Real, R., & Vargas, J. M. (1996). The probabilistic basis of Jaccard's index of similarity. Systematic Biology, 45(3), 380-385. DOI:10.1093/sysbio/45.3.380

- Jaccard < 0.25: Typically considered low similarity, meaning that the two sets have few elements in common.
- 0.25 ≤ Jaccard < 0.5: Moderate similarity, which indicates that the sets share some commonalities but still have many differences.
- 0.5 ≤ Jaccard < 0.75: High similarity, meaning there is substantial overlap between the sets.
- Jaccard ≥ 0.75: Very high similarity, with most elements being shared between the sets.

## Original code

```{r}

##### Estimation of overlap, using the Jaccard Index

# IDS
#dat1<-dat[which(dat$HRSD==1|dat$IDS==1),]
a1 <- 1-(dist.binary(matrix(c(dat$IDS, dat$QIDS), nrow = 2, byrow = TRUE), method = 1)^2)
b1 <- 1-(dist.binary(matrix(c(dat$IDS, dat$BDI), nrow = 2, byrow = TRUE), method = 1)^2)
c1 <- 1-(dist.binary(matrix(c(dat$IDS, dat$CESD), nrow = 2, byrow = TRUE), method = 1)^2)
d1 <- 1-(dist.binary(matrix(c(dat$IDS, dat$SDS), nrow = 2, byrow = TRUE), method = 1)^2)
e1 <- 1-(dist.binary(matrix(c(dat$IDS, dat$MADRS), nrow = 2, byrow = TRUE), method = 1)^2)
f1 <- 1-(dist.binary(matrix(c(dat$IDS, dat$HRSD), nrow = 2, byrow = TRUE), method = 1)^2)
IDS.v <- c(1,a1,b1,c1,d1,e1,f1)

# QIDS
a2 <- 1-(dist.binary(matrix(c(dat$QIDS, dat$IDS), nrow = 2, byrow = TRUE), method = 1)^2)
b2 <- 1-(dist.binary(matrix(c(dat$QIDS, dat$BDI), nrow = 2, byrow = TRUE), method = 1)^2)
c2 <- 1-(dist.binary(matrix(c(dat$QIDS, dat$CESD), nrow = 2, byrow = TRUE), method = 1)^2)
d2 <- 1-(dist.binary(matrix(c(dat$QIDS, dat$SDS), nrow = 2, byrow = TRUE), method = 1)^2)
e2 <- 1-(dist.binary(matrix(c(dat$QIDS, dat$MADRS), nrow = 2, byrow = TRUE), method = 1)^2)
f2 <- 1-(dist.binary(matrix(c(dat$QIDS, dat$HRSD), nrow = 2, byrow = TRUE), method = 1)^2)
QIDS.v <- c(a2,1,b2,c2,d2,e2,f2)

# BDI
a3 <- 1-(dist.binary(matrix(c(dat$BDI, dat$IDS), nrow = 2, byrow = TRUE), method = 1)^2)
b3 <- 1-(dist.binary(matrix(c(dat$BDI, dat$QIDS), nrow = 2, byrow = TRUE), method = 1)^2)
c3 <- 1-(dist.binary(matrix(c(dat$BDI, dat$CESD), nrow = 2, byrow = TRUE), method = 1)^2)
d3 <- 1-(dist.binary(matrix(c(dat$BDI, dat$SDS), nrow = 2, byrow = TRUE), method = 1)^2)
e3 <- 1-(dist.binary(matrix(c(dat$BDI, dat$MADRS), nrow = 2, byrow = TRUE), method = 1)^2)
f3 <- 1-(dist.binary(matrix(c(dat$BDI, dat$HRSD), nrow = 2, byrow = TRUE), method = 1)^2)
BDI.v <- c(a3,b3,1,c3,d3,e3,f3)

# CESD
a4 <- 1-(dist.binary(matrix(c(dat$CESD, dat$IDS), nrow = 2, byrow = TRUE), method = 1)^2)
b4 <- 1-(dist.binary(matrix(c(dat$CESD, dat$QIDS), nrow = 2, byrow = TRUE), method = 1)^2)
c4 <- 1-(dist.binary(matrix(c(dat$CESD, dat$BDI), nrow = 2, byrow = TRUE), method = 1)^2)
d4 <- 1-(dist.binary(matrix(c(dat$CESD, dat$SDS), nrow = 2, byrow = TRUE), method = 1)^2)
e4 <- 1-(dist.binary(matrix(c(dat$CESD, dat$MADRS), nrow = 2, byrow = TRUE), method = 1)^2)
f4 <- 1-(dist.binary(matrix(c(dat$CESD, dat$HRSD), nrow = 2, byrow = TRUE), method = 1)^2)
CESD.v <- c(a4,b4,c4,1, d4,e4,f4)

# SDS
a5 <- 1-(dist.binary(matrix(c(dat$SDS, dat$IDS), nrow = 2, byrow = TRUE), method = 1)^2)
b5 <- 1-(dist.binary(matrix(c(dat$SDS, dat$QIDS), nrow = 2, byrow = TRUE), method = 1)^2)
c5 <- 1-(dist.binary(matrix(c(dat$SDS, dat$BDI), nrow = 2, byrow = TRUE), method = 1)^2)
d5 <- 1-(dist.binary(matrix(c(dat$SDS, dat$CESD), nrow = 2, byrow = TRUE), method = 1)^2)
e5 <- 1-(dist.binary(matrix(c(dat$SDS, dat$MADRS), nrow = 2, byrow = TRUE), method = 1)^2)
f5 <- 1-(dist.binary(matrix(c(dat$SDS, dat$HRSD), nrow = 2, byrow = TRUE), method = 1)^2)
SDS.v <- c(a5,b5,c5, d5,1, e5,f5)

# MADRS
a6 <- 1-(dist.binary(matrix(c(dat$MADRS, dat$IDS), nrow = 2, byrow = TRUE), method = 1)^2)
b6 <- 1-(dist.binary(matrix(c(dat$MADRS, dat$QIDS), nrow = 2, byrow = TRUE), method = 1)^2)
c6 <- 1-(dist.binary(matrix(c(dat$MADRS, dat$BDI), nrow = 2, byrow = TRUE), method = 1)^2)
d6 <- 1-(dist.binary(matrix(c(dat$MADRS, dat$CESD), nrow = 2, byrow = TRUE), method = 1)^2)
e6 <- 1-(dist.binary(matrix(c(dat$MADRS, dat$SDS), nrow = 2, byrow = TRUE), method = 1)^2)
f6 <- 1-(dist.binary(matrix(c(dat$MADRS, dat$HRSD), nrow = 2, byrow = TRUE), method = 1)^2)
MADRS.v <- c(a6,b6,c6, d6,e6,1,f6)

# HRSD
a7 <- 1-(dist.binary(matrix(c(dat$HRSD, dat$IDS), nrow = 2, byrow = TRUE), method = 1)^2)
b7 <- 1-(dist.binary(matrix(c(dat$HRSD, dat$QIDS), nrow = 2, byrow = TRUE), method = 1)^2)
c7 <- 1-(dist.binary(matrix(c(dat$HRSD, dat$BDI), nrow = 2, byrow = TRUE), method = 1)^2)
d7 <- 1-(dist.binary(matrix(c(dat$HRSD, dat$CESD), nrow = 2, byrow = TRUE), method = 1)^2)
e7 <- 1-(dist.binary(matrix(c(dat$HRSD, dat$SDS), nrow = 2, byrow = TRUE), method = 1)^2)
f7 <- 1-(dist.binary(matrix(c(dat$HRSD, dat$MADRS), nrow = 2, byrow = TRUE), method = 1)^2)
HRSD.v <- c(a7,b7,c7, d7,e7,f7,1)

# Create table
M = matrix(nrow=7, ncol=7) 
colnames(M) <- c("IDS",	"QIDS",	"BDI",	"CESD",	"SDS"	,"MADRS"	,"HRSD")
rownames(M) <- c("IDS",	"QIDS",	"BDI",	"CESD",	"SDS"	,"MADRS"	,"HRSD")
M[1,] <- IDS.v
M[2,] <- QIDS.v
M[3,] <- BDI.v
M[4,] <- CESD.v
M[5,] <- SDS.v
M[6,] <- MADRS.v
M[7,] <- HRSD.v
isSymmetric(M)

janitor::round_half_up(M, 2)

M[M == 1] <- 0 # replace diagonal with 0
janitor::round_half_up(colMeans(M, na.rm = TRUE), 2)
janitor::round_half_up(mean(colMeans(M, na.rm = TRUE)), 2) 

```

Eiko replaced the diagonal with 0 and included this in the mean, although the table in the manuscript puts 1 in the diagonal.

"The mean overlap among all scales is 0.36" ([Fried, 2017, p. 194)

“Mean overlap 0.39 0.40 0.40 0.27 0.37 0.30 0.36” (Fried, 2017, p. 194)

### Exclude diagonal from mean calculation

```{r}

diag(M) <- NA
janitor::round_half_up(colMeans(M, na.rm = TRUE), 2)

```

The mean overlap among all scales is `r janitor::round_half_up(mean(colMeans(M, na.rm = TRUE)), 2)`

## Using vegan

```{r}

# calculate Jaccard similarities
jaccard_diff <- dat |>
  as.matrix() |>
  t() |>
  vegan::vegdist(method = "jaccard", binary = TRUE) |>
  as.matrix()

jaccard_sim <- 1 - jaccard_diff

# round
jaccard_sim_rounded <- janitor::round_half_up(jaccard_sim, 2)

# print
jaccard_sim_rounded

# calculate column means
diag(jaccard_sim_rounded) <- NA
janitor::round_half_up(colMeans(jaccard_sim_rounded, na.rm = TRUE), 2)

```

The mean overlap among all scales is `r janitor::round_half_up(mean(colMeans(jaccard_sim_rounded, na.rm = TRUE)), 2)`

## Using proxy

```{r}

# calculate Jaccard similarities
jaccard_diff <- dat |>
  as.matrix() |> 
  t() |>
  proxy::dist(data_mat_t, method = "Jaccard") |>
  as.matrix()

jaccard_sim <- 1 - jaccard_diff

# round
jaccard_sim_rounded <- janitor::round_half_up(jaccard_sim, 2)

# print
jaccard_sim_rounded

# calculate column means
diag(jaccard_sim_rounded) <- NA
janitor::round_half_up(colMeans(jaccard_sim_rounded, na.rm = TRUE), 2)

```

The mean overlap among all scales is `r janitor::round_half_up(mean(colMeans(jaccard_sim_rounded, na.rm = TRUE)), 2)`

# original other calc

```{r}

length1<-c(28,9,21,18,20,9,17) # length of original (adjusted) questionnaires; e.g. CESD 20 items originally, combined down to 18
length2<-c(33,20,25,21,23,12,22) # items in analysis per scale; CESD captures 21 items

cor(length1, colMeans(M)) #0.2938447
cor(length2, colMeans(M)) #0.5720158

```

# original plot

```{r}

dataplot <- read_csv("MatrixA.csv") #Load data for plot (difference between specific and compound symptoms)

##### Figure 1 
##### Thanks for assistance with ggplot2 to Jana Jarecki
set.seed(223)
d <- dataplot
d[, S := factor(paste0("S",1:nrow(d)))] #Create symptom variable
d = melt(d, id.vars="S", variable.name="Scales", value.name="Type") #Transform to long format
d = d[Type>=1] #Keep the scales in which the symptoms are 1 (present) or 2 (included)
d[, Type := factor(Type, labels=c("Scale contains compound symptom", "Scale contains specific symptom"))]
d[, count := .N, by=S]

# Symptom order
sympt.order = d[, .N, by=S][order(N)][, S] #Replace by order
d[, S := factor(S, levels = sympt.order)]

# Scale order by frequency
scale.order = d[, .N, by=Scales][order(N)][, Scales]
d[, Scales := factor(Scales, levels = scale.order)]
d[, Scales2 := as.numeric(Scales)]

# Plot
pal1 <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

a<- ggplot(d, aes(x=S, y=Scales2, group=S, color=as.factor(Scales), shape=Type, rev=F)) +
  geom_line() + #keep this here, otherwise there is an error 
  xlab("") +
  ylab("") +
  # Generate the grid lines
  geom_hline(yintercept = 1:7, colour = "grey80", size = .2) +
  geom_vline(xintercept = 1:52, colour = "grey80", size = .2) +
  # Points and lines
  geom_line(colour="grey60") +
  geom_point(size=3, fill="white") +
  # Fill the middle space with a white blank rectangle
  geom_rect(xmin=-Inf,xmax=Inf,ymin=-Inf,ymax=.6,fill="white", color=NA) +
  # Polar coordinates
  coord_polar() +
  scale_shape_manual(values=c(21,19)) +
  # The angle for the symptoms and remove the default grid lines
  theme(axis.text.x = element_text(angle = 360/(2*pi)*rev( pi/2 + seq( pi/52, 2*pi-pi/52, len=52)) + c(rep(0, 26), rep(180,28))),
        panel.border = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        legend.position="top",
        plot.margin = unit(rep(.5,4), "lines")) +
  labs(fill="") + # remove legend title
  scale_y_continuous(limits=c(-4,7), expand=c(0,0), breaks=1:7, labels=d[, levels(Scales)]) +
  scale_color_manual(values=pal1); a

ggsave(plot=a,filename="Figure1.pdf", width=7, height=7, useDingbats=FALSE)
# Figure 1 was further adjusted in Inkscape

```




